Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
# load data
setwd('~/Desktop/ud651-r-eda/lesson3/')
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

# load ggplot2
library(ggplot2)

# plot scatter
qplot(x = age, y = friend_count, data = pf)

# plot scatter (using ggplot syntax)
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() +
  xlim(13,90)

summary(pf$age)

#

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes: To reduce overplotting, set alpha

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13,90)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + # jitter adds some noise to each age
  xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  coord_trans(y = "sqrt") +
  xlim(13,90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
# friendships initiated vs age
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20)
  #coord_trans(y = "sqrt") +
  #xlim(13,90)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library('dplyr')

age_groups <- group_by(pf, age)

pf.fc_by_age <- summarise(age_groups,
                  friend_count_mean = mean(friend_count),
                  friend_count_median = median(friend_count),
                  n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

# alternative syntax, using function chaining

pf.fc_by_age <- pf %.%
  group_by(age) %.%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age)
```

Create your plot!

```{r Conditional Means Plot}

ggplot(aes(x = age, y = friend_count_median), data = pf.fc_by_age) +
  geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  xlim(13,90) +
  geom_point(alpha = 1/20,
             position = position_jitter(h=0),
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  coord_cartesian(xlim = c(13,90)) +
  geom_point(alpha = 1/20,
             position = position_jitter(h=0),
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = .1,
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .5,
            linetype = 1, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .9,
            linetype = 2, color = 'blue')


```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95))
  #coord_trans(y = 'sqrt', x = 'sqrt') +
  #geom_line(stat = 'summary', fun.y = mean)
  #geom_line(stat = 'summary', fun.y = quantile, probs = .1,
  #          linetype = 2, color = 'blue') +
  #geom_line(stat = 'summary', fun.y = quantile, probs = .5,
  #          linetype = 1, color = 'blue') +
  #geom_line(stat = 'summary', fun.y = quantile, probs = .9,
  #          linetype = 2, color = 'blue')
  
  

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

with(Mitchell, cor.test(Month, Temp))

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf $age + (12 - pf$dob_month) /12
```

Programming Assignment
```{r Programming Assignment}

install.packages('dplyr')
library('dplyr')

pf.fc_by_age_months <- pf %.%
  group_by(age_with_months) %.%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age_with_months)

ggplot(aes(x = age_with_months, y = friend_count_median), data = pf.fc_by_age_months) +
  geom_line()

ggplot(aes(x = age_with_months,
           y = friend_count_mean),
       data = subset(pf.fc_by_age_months, pf.fc_by_age_months$age <= 71)) +
  geom_line()



# show together

# extra noisey (smaller bins)
p1 <- ggplot(aes(x = age_with_months,
           y = friend_count_mean),
       data = subset(pf.fc_by_age_months, pf.fc_by_age_months$age < 71)) +
  geom_line()

# original
p2 <- ggplot(aes(x = age,
                 y = friend_count_mean),
       data = subset(pf.fc_by_age, pf.fc_by_age$age < 71)) +
  geom_line()

# extra smooth (wider bins - rounded to nearest 5)
p3 <- ggplot(aes(x = round(age / 5) * 5,
                 y = friend_count),
       data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol=1)

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
# extra noisey (smaller bins)
p1 <- ggplot(aes(x = age_with_months,
           y = friend_count_mean),
       data = subset(pf.fc_by_age_months, pf.fc_by_age_months$age < 71)) +
  geom_line() +
  geom_smooth()

# original
p2 <- ggplot(aes(x = age,
                 y = friend_count_mean),
       data = subset(pf.fc_by_age, pf.fc_by_age$age < 71)) +
  geom_line()+
  geom_smooth()

# extra smooth (wider bins - rounded to nearest 5)
p3 <- ggplot(aes(x = round(age / 5) * 5,
                 y = friend_count),
       data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

